---
title: "STA 9750 Mini-Project #01: Netflix Analysis"
author: "Eduardo Alarcon"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
execute:
  message: false
  warning: false
---

```{r setup}
library(dplyr)
library(lubridate)
library(readr)
library(scales)
library(tidyverse)
library(knitr)
```

```{r}
#| include: false 

if(!dir.exists(file.path("data", "mp01"))){ dir.create(file.path("data", "mp01"), showWarnings=FALSE, recursive=TRUE) }
GLOBAL_TOP_10_FILENAME <- file.path("data", "mp01", "global_top10_alltime.csv")
if(!file.exists(GLOBAL_TOP_10_FILENAME)){ download.file("https://www.netflix.com/tudum/top10/data/all-weeks-global.tsv", destfile=GLOBAL_TOP_10_FILENAME) }
COUNTRY_TOP_10_FILENAME <- file.path("data", "mp01", "country_top10_alltime.csv")
if(!file.exists(COUNTRY_TOP_10_FILENAME)){ download.file("https://www.netflix.com/tudum/top10/data/all-weeks-countries.tsv", destfile=COUNTRY_TOP_10_FILENAME) }

if(!require("tidyverse")) install.packages("tidyverse")

GLOBAL_TOP_10 <- read_tsv(GLOBAL_TOP_10_FILENAME)
str(GLOBAL_TOP_10)
glimpse(GLOBAL_TOP_10)

GLOBAL_TOP_10 <- GLOBAL_TOP_10 %>%
  mutate(season_title = if_else(season_title == "N/A", NA, season_title))
glimpse(GLOBAL_TOP_10)

COUNTRY_TOP_10 <- read_tsv(COUNTRY_TOP_10_FILENAME, na = "N/A")
glimpse(COUNTRY_TOP_10)
str(COUNTRY_TOP_10)

n_distinct(COUNTRY_TOP_10$country_name)

GLOBAL_TOP_10 %>%
filter(category == "Films (Non-English)") %>%
group_by(show_title) %>%
summarise(max_cumulweeks = max(cumulative_weeks_in_top_10, na.rm = TRUE)) %>%
arrange(desc(max_cumulweeks)) %>%
slice(1)

library(stringr)
library(DT)

format_titles <- function(df){
    colnames(df) <- str_replace_all(colnames(df), "_", " ") |> str_to_title()
    df
}

GLOBAL_TOP_10 |> 
head(n=20) |>
datatable(options=list(searching=FALSE, info=FALSE))

GLOBAL_TOP_10 |> 
format_titles() |>
head(n=20) |>
datatable(options=list(searching=FALSE, info=FALSE)) |>
formatRound(c('Weekly Hours Viewed', 'Weekly Views'))

```

## Task 4

The following questions address Task 4. Please click on the "Show Code" for an detailed approach on how I obtained my results.

<br>

1. How many different countries does Netflix operate in? (You can use the viewing history as a proxy for countries in which Netflix operates.)

```{r}
#| output: false
n_distinct(COUNTRY_TOP_10$country_name)
``` 
**Answer:** Netflix operates in ```r n_distinct(COUNTRY_TOP_10$country_name)``` different countries based on the viewing history data.

<br>

2. Which non-English-language film has spent the most cumulative weeks in the global top 10? How many weeks did it spend?

```{r}
Q2_RESULT <- GLOBAL_TOP_10 %>%
filter(category == "Films (Non-English)") %>%
group_by(show_title) %>%
summarise(`Maximum Cumulative Weeks` = max(cumulative_weeks_in_top_10, na.rm = TRUE)) %>%
arrange(desc(`Maximum Cumulative Weeks`)) %>%
slice(1)

film_name_2 <- Q2_RESULT$show_title
weeks_count_2 <- Q2_RESULT[[2]]

Q2_RESULT %>%
format_titles() %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))

```

<br>

**Answer:** The non-English-language film `r film_name_2` spent the most cumulative weeks in the Global Top 10, with `r weeks_count_2` weeks.

<br>

3.	What is the longest film (English or non-English) to have ever appeared in the Netflix global Top 10? How long is it in minutes?
```{r}
Q3_RESULT <- GLOBAL_TOP_10 %>%
  filter(category %in% c("Films (Non-English)", "Films (English)")) %>%
  filter(!is.na(runtime)) %>%   
  group_by(show_title) %>%
  summarise(max_runtime = max(runtime)) %>%   
  arrange(desc(max_runtime)) %>%
  slice(1) %>%
  mutate(max_runtime_minutes = round(60 * max_runtime))

film_name_3 <- Q3_RESULT$show_title
run_time_3 <- Q3_RESULT[[3]]

Q3_RESULT %>%
select(-max_runtime) %>%
format_titles() %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))

```

<br>

**Answer:** The longest English or non-English film to have ever topped in the Netflix Global Top 10 is `r film_name_3`, with a length of `r run_time_3` minutes.

<br>

4. For each of the four categories, what program has the most total hours of global viewership?

```{r}
Q4_RESULT <- GLOBAL_TOP_10 %>%
group_by(category, show_title) %>%
summarise(total_hours_global_viewership = sum(weekly_hours_viewed, na.rm = TRUE), .groups = "drop") %>%
group_by(category) %>%
slice_max(total_hours_global_viewership, n = 1) %>%
arrange(desc(total_hours_global_viewership)) %>%
mutate(total_hours_global_viewership = comma(total_hours_global_viewership))

Q4_RESULT %>%
format_titles() %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

**Answer:** The above table highlights calculations that sum all seasons together (i.e.,franchise-level data). In other words, Stranger Things and Squid Game reflect the combined global viewership across all of their respective seasons.

<br>

5.	Which TV show had the longest run in a country’s Top 10? How long was this run and in what country did it occur?

```{r}

Q5_RESULT <- COUNTRY_TOP_10 %>%
filter(category == "TV") %>%
filter(!is.na(season_title)) %>%
group_by(country_name, show_title) %>%
summarise(max_cumulative_weeks = max(cumulative_weeks_in_top_10), .groups = "drop") %>%
arrange(desc(max_cumulative_weeks)) %>%
slice(1)

Q5_RESULT %>%
format_titles() %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

**Answer:** Money Heist had the longest run in Pakistan, with 127 weeks in Netflix's Country Top 10 list. 

<br>

6. Netflix provides over 200 weeks of service history for all but one country in our data set. Which country is this and when did Netflix cease operations in that country?

```{r}

Q6_RESULT <- COUNTRY_TOP_10 %>%
group_by(country_name) %>%
summarise(total_weeks_of_data = n_distinct(week),final_week = max(week, na.rm = TRUE)) %>%
filter(total_weeks_of_data < 200) %>%
arrange(total_weeks_of_data)

Q6_RESULT %>%
format_titles() %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))

```

<br>

**Answer:** Netflix could not provide over 200 weeks of service for Russia, as it ceased operations during the week of February 27, 2022, resulting in 35 weeks of data in the Country Top 10 list.

<br>

7. What is the total viewership of the TV show Squid Game? Note that there are three seasons total and we are looking for the total number of hours watched across all seasons.

```{r}
Q7_RESULT_SEASONS <- GLOBAL_TOP_10 %>%
filter(show_title == "Squid Game") %>%
filter(!is.na(season_title)) %>%
group_by(season_title) %>%
summarise(sum_hours = sum(weekly_hours_viewed, na.rm = TRUE)) %>%
mutate(hours_watched = comma(sum_hours))

Q7_RESULT_SEASONS_SUM <- Q7_RESULT_SEASONS %>%
summarise(
season_title = "Total Hours Watched Across All Seasons",
sum_hours_1 = sum(sum_hours),
hours_watched = comma(sum_hours_1))

Q7_RESULT_FINAL <- bind_rows(Q7_RESULT_SEASONS, Q7_RESULT_SEASONS_SUM) %>%
select(season_title, hours_watched)

Q7_RESULT_FINAL %>%
format_titles() %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

**Answer:** Across all three seasons, Squid Game had a total of 5,048,300,000 hours of global viewership. 

<br>

8. The movie Red Notice has a runtime of 1 hour and 58 minutes. Approximately how many views did it receive in 2021?

```{r}
Q8_RESULT <- GLOBAL_TOP_10 %>%
filter(show_title == "Red Notice") %>%
filter(year(week) == 2021) %>%
summarise(
total_hours_2021 = sum(weekly_hours_viewed, na.rm = TRUE),
runtime_hours = 1 + 58/60,
approximate_views = total_hours_2021 / runtime_hours) %>%
mutate(
approximate_views_formatted = comma(round(approximate_views)),
total_hours_formatted = comma(total_hours_2021)
)

Q8_RESULT %>%
select(total_hours_formatted, approximate_views_formatted) %>%
rename(
`Total Hours Viewed (2021)` = total_hours_formatted,
`Approximate Views` = approximate_views_formatted) %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

**Answer:** Based on the total number of hours viewed in 2021, Red Notice has approximately, 201,732,203 global views.

<br>

9. Part A: How many Films reached Number 1 in the US but did not originally debut there? 


```{r}

COUNTRY_TOP_10_US_ONLY <- COUNTRY_TOP_10 %>%
  filter(country_iso2 == "US", category == "Films") %>%
  mutate(week = as.Date(week))

US_FILMS_REACH1 <- COUNTRY_TOP_10_US_ONLY %>%
  group_by(show_title) %>%
  summarise(best_rank = min(weekly_rank, na.rm = TRUE), .groups = "drop") %>%
  filter(best_rank == 1)

US_DEBUT <- COUNTRY_TOP_10_US_ONLY %>%
  group_by(show_title) %>%
  arrange(week, .by_group = TRUE) %>%   # <-- critical: order before taking first
  slice(1L) %>%                         # earliest appearance in US
  ungroup() %>%
  transmute(show_title, debut_week = week, debut_rank = weekly_rank)

Q9_RESULT_PT1 <- US_FILMS_REACH1 %>%
  inner_join(US_DEBUT, by = "show_title") %>%
  filter(debut_rank > 1) %>%
  summarise(number_of_films = n())

```

<br>

**Answer:** A total of ```r Q9_RESULT_PT1``` films in the US reached to Number 1 after debuting at a lower ranking. 

<br>

9. Part B: What is the most recent film to pull this off?

```{r}

Q9_RESULT_PT2 <- COUNTRY_TOP_10 %>%
filter(country_iso2 == "US", category == "Films") %>%
mutate(week = as.Date(week)) %>%
group_by(show_title) %>%
summarise(
debut_week = min(week),
debut_rank = weekly_rank[which.min(week)],
best_rank  = min(weekly_rank, na.rm = TRUE),
.groups = "drop") %>%
filter(best_rank == 1, debut_rank > 1) %>%
inner_join(COUNTRY_TOP_10 %>%
filter(country_iso2 == "US", category == "Films") %>%
mutate(week = as.Date(week)) %>%
select(show_title, week, weekly_rank), by = "show_title") %>%
filter(weekly_rank == 1) %>%
slice_max(order_by = week, n = 1, with_ties = FALSE) %>%
transmute(`Most Recent Film` = show_title, `Date Reached #1` = week)

# Present as datatable
Q9_RESULT_PT2 %>%
datatable(
  options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))

```

<br>

**Answer:** The most recent film to accomplish this was KPop Demon Hunters, which reached Number 1 during the week of September 14, 2025. 

<br>

10. Which TV show/season hit the top 10 in the most countries in its debut week? In how many countries did it chart?

```{r}

COUNTRY_TOP_10_DEBUT <- COUNTRY_TOP_10 %>%
filter(category == "TV") %>%
group_by(show_title, season_title, country_name) %>%
summarise(
debut_week = min(week, na.rm = TRUE),
debut_rank = weekly_rank[which.min(week)][1],
.groups = "drop"
) %>%
arrange(debut_week)

Q10_RESULT <- COUNTRY_TOP_10_DEBUT %>% 
group_by(show_title, season_title, debut_week) %>% 
summarise(country_appearance = n_distinct(country_name), .groups = "drop") %>% 
arrange(desc(country_appearance)) %>% 
slice(1) %>%
rename(`Show Title` = show_title, `Season Title` = season_title, `Debut Week` = debut_week, `Number of Countries Charted` = country_appearance)

Q10_RESULT %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

**Answer:** Emily in Paris (Season 2) charted in 94 countries during its debut week in December 26, 2021. 

<br>

## Task 5: 
## Press Release 1: Stranger Things

The following task involves creating a "Press Release" for the upcoming release of the final season of Stranger Things.  

<br>

```{r}
STRANGE_GLOBAL_VIEWERSHIP <- GLOBAL_TOP_10 %>%
filter(show_title == "Stranger Things") %>%
mutate(season_display = ifelse(is.na(season_title), "Stranger Things: S1", season_title)) %>%
group_by(season_display) %>%
summarise(total_hours = sum(weekly_hours_viewed, na.rm = TRUE)) %>%
mutate(season_display = case_when(
season_display == "Stranger Things 2" ~ "Stranger Things: S2",
season_display == "Stranger Things 3" ~ "Stranger Things: S3",
season_display == "Stranger Things 4" ~ "Stranger Things: S4", TRUE ~ season_display)) %>%
mutate(avg_episode_minutes = case_when(
season_display == "Stranger Things: S1" ~ 47,
season_display == "Stranger Things: S2" ~ 60,
season_display == "Stranger Things: S3" ~ 60,
season_display == "Stranger Things: S4" ~ 87, TRUE ~ NA_real_),
avg_episode_hours = avg_episode_minutes / 60,
approximate_views = total_hours / avg_episode_hours,
total_hours_viewed = comma(total_hours),
approximate_viewership = comma(round(approximate_views))) %>%
arrange(season_display)

STRANGE_TOTAL_VIEWERSHIP <- STRANGE_GLOBAL_VIEWERSHIP %>%
filter(!is.na(avg_episode_minutes)) %>%
summarise(
season_display = "Totals",
total_hours = sum(total_hours),
total_views = sum(approximate_views, na.rm = TRUE),
avg_episode_minutes = NA_real_,
total_hours_viewed = comma(total_hours),
approximate_viewership = comma(round(total_views)))

STRANGE_VIEWERSHIP_SYNTH <- bind_rows(
STRANGE_GLOBAL_VIEWERSHIP %>% select(season_display, total_hours_viewed, avg_episode_minutes, approximate_viewership),
STRANGE_TOTAL_VIEWERSHIP %>% select(season_display, total_hours_viewed, avg_episode_minutes, approximate_viewership)
) %>%
rename( Season = season_display, `*Average Episode Minutes` = avg_episode_minutes)

STRANGE_VIEWERSHIP_SYNTH %>%
format_titles() %>%
datatable(
options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

*Average minutes derived from the following sources:

*[Wikipedia's List of Stranger Things Episodes](https://en.wikipedia.org/wiki/List_of_Stranger_Things_episodes)*

*[Netflix's Official Tudum Blog](https://www.netflix.com/tudum/articles/stranger-things-season-4-episode-length)*

```{r}

STRANGE_POPULARITY <- GLOBAL_TOP_10 %>%
filter(show_title == "Stranger Things") %>%
mutate(season_display = ifelse(is.na(season_title), "Stranger Things: S1", season_title)) %>%
mutate(season_display = case_when(
season_display == "Stranger Things 2" ~ "Stranger Things: S2",
season_display == "Stranger Things 3" ~ "Stranger Things: S3",
season_display == "Stranger Things 4" ~ "Stranger Things: S4",
TRUE ~ season_display)) %>%
group_by(season_display) %>%
summarise(
weeks_in_top_10 = n_distinct(week),
weeks_at_number_1 = sum(weekly_rank == 1, na.rm = TRUE) ) %>%
arrange(season_display) %>%
rename(Season = season_display)

STRANGE_POPULARITY %>%
format_titles() %>%
datatable(
options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

```{r}
STRANGE_COUNTRY10_SUMMARY <- COUNTRY_TOP_10 %>%
filter(show_title == "Stranger Things") %>%
mutate(season_display = ifelse(is.na(season_title), "Stranger Things: S1", season_title)) %>%
group_by(country_name) %>%
summarise(total_weeks_in_top_10 = n_distinct(week)) %>%
summarise(
`Total Countries` = n(),
`Average Weeks in Top 10` = round(mean(total_weeks_in_top_10), 1))

STRANGE_COUNTRY10_SUMMARY %>%
datatable(
options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

```{r}
ST_TOP_COUNTRIES_WEEKS <- COUNTRY_TOP_10 %>%
filter(show_title == "Stranger Things") %>%
mutate(season_display = ifelse(is.na(season_title), "Stranger Things: S1", season_title)) %>%
group_by(country_name) %>%
summarise(total_weeks_in_top_10 = n_distinct(week)) %>%
arrange(desc(total_weeks_in_top_10)) %>%
slice(1:10) %>%
rename(Country = country_name)

ST_TOP_COUNTRIES_WEEKS %>%
format_titles() %>%
datatable(
options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))
```

<br>

## Task 6:
## Press Release 2: Commericial Success in India

<br>

```{r}
# Get the detailed list first
INDIA_SHOWS_LIST <- COUNTRY_TOP_10 %>%
filter(country_name == "India") %>%
distinct(show_title, season_title, category) %>%
mutate(content_type = case_when(
str_detect(category, "Films") ~ "Films",
str_detect(category, "TV") ~ "TV Shows",TRUE ~ "Other")) %>%
arrange(content_type, show_title, season_title)

# Get the summary counts
INDIA_TOTAL_PROGRAMS <- INDIA_SHOWS_LIST %>%
count(content_type, name = "distinct_shows") %>%
mutate(percentage = round(distinct_shows / sum(distinct_shows) * 100, 1))

# Display the summary table
INDIA_TOTAL_PROGRAMS %>%
rename(`Content Type` = content_type, `Number of Shows` = distinct_shows, `Percentage` = percentage) %>%
datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))

# Show total count
total_shows <- sum(INDIA_TOTAL_PROGRAMS$distinct_shows)
cat("Total distinct viewing options in India:", total_shows)
```

<br>

```{r}

# Get shows that appear in India but NOT in the United States (considering show+season combinations)
india_exclusive_from_us <- COUNTRY_TOP_10 %>%
  group_by(show_title, season_title) %>%
  summarise(
    appears_in_india = any(country_name == "India"),
    appears_in_us = any(country_name == "United States"),
    .groups = "drop"
  ) %>%
  filter(appears_in_india == TRUE & appears_in_us == FALSE)

# Get the detailed list first, excluding Telugu and Tamil content
india_only_list <- COUNTRY_TOP_10 %>%
  filter(country_name == "India") %>%
  semi_join(india_exclusive_from_us, by = c("show_title", "season_title")) %>%
  filter(!str_detect(tolower(show_title), "telugu|tamil")) %>%
  distinct(show_title, season_title, category) %>%
  mutate(content_type = case_when(
    str_detect(category, "Films") ~ "Films",
    str_detect(category, "TV") ~ "TV Shows",
    TRUE ~ "Other"
  )) %>%
  arrange(content_type, show_title, season_title)

# Get breakdown by content type
india_only_breakdown <- india_only_list %>%
  count(content_type, name = "distinct_shows") %>%
  mutate(percentage = round(distinct_shows / sum(distinct_shows) * 100, 1))

# Display the summary results
india_only_breakdown %>%
  rename(`Content Type` = content_type, `Number of Shows` = distinct_shows, `Percentage` = percentage) %>%
  datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))))

# Show total count
total_india_only <- sum(india_only_breakdown$distinct_shows)
cat("Total distinct viewing options in India that do NOT 
  appear in US (excluding Telugu/Tamil language viewing options):", total_india_only)

```

<br>

```{r}

# Get India-exclusive show+season combinations (excluding Telugu/Tamil)
india_exclusive_combinations <- COUNTRY_TOP_10 %>%
  group_by(show_title, season_title) %>%
  summarise(
    appears_in_india = any(country_name == "India"),
    appears_in_us = any(country_name == "United States"),
    .groups = "drop"
  ) %>%
  filter(appears_in_india == TRUE & appears_in_us == FALSE) %>%
  filter(!str_detect(tolower(show_title), "telugu|tamil"))

# Table 1: India-exclusive NON-ENGLISH content that achieved global top 10 status
global_appearing_breakdown <- GLOBAL_TOP_10 %>%
  filter(category %in% c("Films (Non-English)", "TV (Non-English)")) %>%
  semi_join(india_exclusive_combinations, by = c("show_title", "season_title")) %>%
  distinct(show_title, season_title, category) %>%
  mutate(content_type = case_when(
    str_detect(category, "Films") ~ "Films",
    str_detect(category, "TV") ~ "TV Shows",
    TRUE ~ "Other"
  )) %>%
  count(content_type, name = "distinct_shows") %>%
  mutate(percentage = round(distinct_shows / sum(distinct_shows) * 100, 1))

# Calculate separate metrics for Films and TV Shows (hours only)
films_metrics <- GLOBAL_TOP_10 %>%
  filter(category == "Films (Non-English)") %>%
  semi_join(india_exclusive_combinations, by = c("show_title", "season_title")) %>%
  summarise(total_hours = sum(weekly_hours_viewed, na.rm = TRUE))

tv_metrics <- GLOBAL_TOP_10 %>%
  filter(category == "TV (Non-English)") %>%
  semi_join(india_exclusive_combinations, by = c("show_title", "season_title")) %>%
  summarise(total_hours = sum(weekly_hours_viewed, na.rm = TRUE))

# Calculate overall summary metrics for NON-ENGLISH content
shows_in_global_count <- sum(global_appearing_breakdown$distinct_shows)
total_combinations <- nrow(india_exclusive_combinations)
global_penetration <- round((shows_in_global_count / total_combinations) * 100, 1)

INDIA_EXCLUSIVE_SUMMARY <- GLOBAL_TOP_10 %>%
  filter(category %in% c("Films (Non-English)", "TV (Non-English)")) %>%
  semi_join(india_exclusive_combinations, by = c("show_title", "season_title")) %>%
  summarise(
    total_hours_viewed = sum(weekly_hours_viewed, na.rm = TRUE),
    avg_weeks_in_top_10 = mean(cumulative_weeks_in_top_10, na.rm = TRUE),
    max_weeks_in_top_10 = max(cumulative_weeks_in_top_10, na.rm = TRUE)
  ) %>%
  mutate(
    total_hours_formatted = comma(total_hours_viewed),
    avg_weeks_formatted = round(avg_weeks_in_top_10, 1)
  )

# Create comprehensive summary table (hours only)
GLOBAL_REACH_SUMMARY <- data.frame(
  Metric = c("Total Show+Season Combinations", 
             "Total Hours Viewed - Non-English Films", 
             "Total Hours Viewed - Non-English TV",
             "Combined Hours Viewed (Non-English)", 
             "Average Weeks in Global Top 10", "Maximum Weeks in Global Top 10", 
             "Non-English Combinations in Global Charts", "Non-English Global Chart Penetration Rate"),
  Value = c(
    total_combinations,
    comma(films_metrics$total_hours), 
    comma(tv_metrics$total_hours),
    INDIA_EXCLUSIVE_SUMMARY$total_hours_formatted,
    paste0(INDIA_EXCLUSIVE_SUMMARY$avg_weeks_formatted, " weeks"),
    paste0(INDIA_EXCLUSIVE_SUMMARY$max_weeks_in_top_10, " weeks"),
    paste0(shows_in_global_count, " of ", total_combinations, " combinations"),
    paste0(global_penetration, "%")
  )
)

# Display India-exclusive NON-ENGLISH content that achieved global status
cat("India-exclusive NON-ENGLISH content that achieved global top 10 status:\n")
global_appearing_breakdown %>%
  rename(`Content Type` = content_type, `Number of Shows` = distinct_shows, `Percentage` = percentage) %>%
  datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE))

cat("Total NON-ENGLISH appearing in global charts:", shows_in_global_count, "of", total_combinations, "\n\n")

# Display comprehensive metrics table (hours only)
GLOBAL_REACH_SUMMARY %>%
  datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE))
```


<br>


```{r}

# Get show+season combinations that appear in India but NOT in the United States
india_exclusive_combinations <- COUNTRY_TOP_10 %>%
  group_by(show_title, season_title) %>%
  summarise(
    appears_in_india = any(country_name == "India"),
    appears_in_us = any(country_name == "United States"),
    .groups = "drop"
  ) %>%
  filter(appears_in_india == TRUE & appears_in_us == FALSE) %>%
  filter(!str_detect(tolower(show_title), "telugu|tamil"))

# Cross-reference with GLOBAL_TOP_10 - FILTER FOR NON-ENGLISH ONLY
INDIA_EXCLUSIVE_METRICS <- GLOBAL_TOP_10 %>%
  filter(category %in% c("Films (Non-English)", "TV (Non-English)")) %>%  # ADD THIS LINE
  semi_join(india_exclusive_combinations, by = c("show_title", "season_title")) %>%
  group_by(show_title, season_title, category) %>%
  summarise(
    total_weekly_hours_viewed = sum(weekly_hours_viewed, na.rm = TRUE),
    avg_runtime_minutes = round(mean(runtime, na.rm = TRUE) * 60, 0),
    total_weekly_views = sum(weekly_views, na.rm = TRUE),
    max_cumulative_weeks_in_top_10 = max(cumulative_weeks_in_top_10, na.rm = TRUE),
    weeks_appeared = n_distinct(week),
    .groups = "drop"
  ) %>%
  mutate(
    content_type = case_when(
      str_detect(category, "Films") ~ "Films",
      str_detect(category, "TV") ~ "TV Shows",
      TRUE ~ "Other"
    ),
    total_hours_formatted = comma(total_weekly_hours_viewed),
    total_views_formatted = comma(total_weekly_views)
  ) %>%
  arrange(desc(total_weekly_hours_viewed)) %>%
  select(show_title, season_title, content_type, total_hours_formatted, avg_runtime_minutes, 
         total_views_formatted, max_cumulative_weeks_in_top_10, weeks_appeared)

# Display the metrics table
INDIA_EXCLUSIVE_METRICS %>%
  format_titles() %>%
  datatable(options = list(searching = TRUE, info = TRUE, paging = TRUE, pageLength = 5
  ))

```

<br>

```{r}

# Get India-exclusive show+season combinations (excluding Telugu/Tamil)
india_exclusive_combinations <- COUNTRY_TOP_10 %>%
  group_by(show_title, season_title) %>%
  summarise(
    appears_in_india = any(country_name == "India"),
    appears_in_us = any(country_name == "United States"),
    .groups = "drop"
  ) %>%
  filter(appears_in_india == TRUE & appears_in_us == FALSE) %>%
  filter(!str_detect(tolower(show_title), "telugu|tamil"))

# Calculate yearly totals from July 4, 2021 to July 4, 2025
yearly_hours <- GLOBAL_TOP_10 %>%
  filter(category %in% c("Films (Non-English)", "TV (Non-English)")) %>%
  semi_join(india_exclusive_combinations, by = c("show_title", "season_title")) %>%
  mutate(
    year_period = case_when(
      week >= as.Date("2021-07-04") & week < as.Date("2022-07-04") ~ "2021-2022",
      week >= as.Date("2022-07-04") & week < as.Date("2023-07-04") ~ "2022-2023", 
      week >= as.Date("2023-07-04") & week < as.Date("2024-07-04") ~ "2023-2024",
      week >= as.Date("2024-07-04") & week < as.Date("2025-07-04") ~ "2024-2025",
      TRUE ~ "Outside Period"
    )
  ) %>%
  filter(year_period != "Outside Period") %>%
  group_by(year_period) %>%
  summarise(
    total_hours_viewed = sum(weekly_hours_viewed, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(year_period) %>%
  mutate(
    total_hours_formatted = comma(total_hours_viewed),
    yoy_change = ifelse(row_number() == 1, NA, 
                       round(((total_hours_viewed - lag(total_hours_viewed)) / lag(total_hours_viewed)) * 100, 1)),
    yoy_change_formatted = ifelse(is.na(yoy_change), "N/A", 
                                 paste0(ifelse(yoy_change > 0, "+", ""), yoy_change, "%"))
  )

# Create yearly analysis table
YEARLY_ANALYSIS <- yearly_hours %>%
  select(year_period, total_hours_formatted, yoy_change_formatted) %>%
  rename(
    `Year Period` = year_period,
    `Total Hours Viewed` = total_hours_formatted,
    `Year-over-Year Change` = yoy_change_formatted
  )

# Display the yearly analysis
cat("India-exclusive Non-English Content - Yearly Viewing Hours Analysis:\n")
YEARLY_ANALYSIS %>%
  datatable(options = list(searching = FALSE, info = FALSE, paging = FALSE))

# Calculate overall growth metrics
first_year <- yearly_hours$total_hours_viewed[1]
last_year <- yearly_hours$total_hours_viewed[nrow(yearly_hours)]
total_growth <- round(((last_year - first_year) / first_year) * 100, 1)
avg_annual_growth <- round(mean(yearly_hours$yoy_change, na.rm = TRUE), 1)

cat("\nGrowth Summary:\n")
cat("Total Growth (2021-2022 to 2024-2025):", total_growth, "%\n")
cat("Average Annual Growth Rate:", avg_annual_growth, "%")

```

<br>

```{r}

# Define South Asian countries (excluding Nepal)
south_asian_countries <- c("India", "Pakistan", "Bangladesh", "Sri Lanka")

# Calculate unique programs by year for each country
yearly_unique_programs <- COUNTRY_TOP_10 %>%
  filter(country_name %in% south_asian_countries) %>%
  mutate(
    year_period = case_when(
      week >= as.Date("2021-07-04") & week < as.Date("2022-07-04") ~ "2021-2022",
      week >= as.Date("2022-07-04") & week < as.Date("2023-07-04") ~ "2022-2023", 
      week >= as.Date("2023-07-04") & week < as.Date("2024-07-04") ~ "2023-2024",
      week >= as.Date("2024-07-04") & week <= as.Date("2025-07-05") ~ "2024-2025",
      TRUE ~ "Outside Period"
    )
  ) %>%
  filter(year_period != "Outside Period") %>%
  mutate(unique_program = paste(show_title, ifelse(is.na(season_title), "NA_SEASON", season_title), sep = " || ")) %>%
  group_by(country_name, year_period) %>%
  summarise(unique_programs_count = n_distinct(unique_program), .groups = "drop") %>%
  arrange(country_name, year_period)

# Calculate year-over-year changes
yearly_programs_with_change <- yearly_unique_programs %>%
  arrange(country_name, year_period) %>%
  group_by(country_name) %>%
  mutate(
    yoy_change = ifelse(row_number() == 1, NA, 
                       round(((unique_programs_count - lag(unique_programs_count)) / lag(unique_programs_count)) * 100, 1)),
    yoy_change_formatted = ifelse(is.na(yoy_change), "N/A", 
                                 paste0(ifelse(yoy_change > 0, "+", ""), yoy_change, "%"))
  ) %>%
  ungroup()

# Create separate tables for counts and changes
counts_table <- yearly_programs_with_change %>%
  select(country_name, year_period, unique_programs_count) %>%
  pivot_wider(names_from = country_name, values_from = unique_programs_count, values_fill = 0) %>%
  arrange(year_period)

changes_table <- yearly_programs_with_change %>%
  select(country_name, year_period, yoy_change_formatted) %>%
  pivot_wider(names_from = country_name, values_from = yoy_change_formatted, values_fill = "N/A") %>%
  arrange(year_period)

# Combine tables with alternating count and change columns
SOUTH_ASIA_DETAILED_ANALYSIS <- counts_table %>%
  rename(`Year Period` = year_period) %>%
  left_join(changes_table %>% rename(`Year Period` = year_period), by = "Year Period", suffix = c("_Count", "_Change")) %>%
  select(`Year Period`, 
         India_Count, India_Change,
         Pakistan_Count, Pakistan_Change,
         Bangladesh_Count, Bangladesh_Change,
         `Sri Lanka_Count`, `Sri Lanka_Change`) %>%
  rename(
    `India Count` = India_Count,
    `India YoY Change` = India_Change,
    `Pakistan Count` = Pakistan_Count,
    `Pakistan YoY Change` = Pakistan_Change,
    `Bangladesh Count` = Bangladesh_Count,
    `Bangladesh YoY Change` = Bangladesh_Change,
    `Sri Lanka Count` = `Sri Lanka_Count`,
    `Sri Lanka YoY Change` = `Sri Lanka_Change`
  )

# Display the main analysis
cat("South Asian Countries - Unique Programs with Separate Rate of Change Columns:\n")
SOUTH_ASIA_DETAILED_ANALYSIS %>%
  datatable(options = list(
    searching = FALSE, 
    info = FALSE, 
    paging = FALSE,
    scrollX = TRUE,
    columnDefs = list(
      list(className = 'dt-center', targets = "_all"),
      list(width = '80px', targets = 1:8)
    )
  ))

# Calculate summary statistics (excluding 2021-2022)
summary_stats <- yearly_programs_with_change %>%
  filter(year_period != "2021-2022") %>%  # Exclude baseline year
  filter(!is.na(yoy_change)) %>%
  group_by(country_name) %>%
  summarise(
    avg_yoy_change = round(mean(yoy_change, na.rm = TRUE), 1),
    total_growth = round(sum(yoy_change, na.rm = TRUE), 1),
    max_single_year_growth = round(max(yoy_change, na.rm = TRUE), 1),
    years_with_data = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_yoy_change))

# Display summary chart

# Display summary chart - Average YoY Change Only
cat("\n\nSummary: Average Year-over-Year Change by Country (Excluding 2021-2022 Baseline):\n")
summary_stats <- yearly_programs_with_change %>%
filter(year_period != "2021-2022") %>%  # Exclude baseline year
filter(!is.na(yoy_change)) %>%
group_by(country_name) %>%
summarise(
avg_yoy_change = round(mean(yoy_change, na.rm = TRUE), 1),
.groups = "drop") %>%
arrange(desc(avg_yoy_change))  # Back to descending order

summary_stats %>%
rename(
Country = country_name,
`Average YoY Change` = avg_yoy_change) %>%
datatable(options = list(
searching = FALSE, 
info = FALSE, 
paging = FALSE,
columnDefs = list(list(className = 'dt-center', targets = "_all"))))

# Identify the top performer
top_country <- summary_stats$country_name[1]
top_avg_growth <- summary_stats$avg_yoy_change[1]

cat("\nHighlight: ", top_country, " shows the greatest average year-over-year growth rate at ", top_avg_growth, "% annually.")

```

<br>

## Task 7:
## Press Release 3: Commericial Success in India

<br>

```{r}

# Get top 10 Latin American countries (excluding Brazil) and combine with global data
combined_data <- COUNTRY_TOP_10 %>%
  filter(
    country_name %in% c("Colombia", "El Salvador", "Honduras", 
                       "Nicaragua", "Venezuela", "Guatemala", "Ecuador", 
                       "Paraguay", "Panama", "Costa Rica", "Peru", 
                       "Bolivia", "Argentina", "Chile") |
    grepl("Dominican Rep", country_name, ignore.case = TRUE)
  ) %>%
  group_by(country_name) %>%
  slice_max(cumulative_weeks_in_top_10, n = 1) %>%
  arrange(desc(cumulative_weeks_in_top_10)) %>%
  head(10) %>%
  mutate(
    season_title = ifelse(is.na(season_title), "N/A", as.character(season_title))
  ) %>%
  left_join(
    GLOBAL_TOP_10 %>%
      mutate(season_title = ifelse(is.na(season_title), "N/A", as.character(season_title))) %>%
      group_by(show_title, season_title) %>%
      slice_max(week, n = 1),
    by = c("show_title", "season_title"),
    suffix = c("_country", "_global")
  ) %>%
  select(
    Country = country_name,
    `Show Title` = show_title,
    `Season Title` = season_title,
    `Cumulative Weeks in Top 10` = cumulative_weeks_in_top_10_country,
    `Global Hours Viewed` = weekly_hours_viewed,
    `Global Rank` = weekly_rank_global
  )

# Create DataTable
datatable(combined_data,
          caption = "Top Latin American Shows: Regional vs Global Performance",
          options = list(searching = FALSE, info = FALSE, paging = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all"))),
          rownames = FALSE) %>%
  formatCurrency('Global Hours Viewed', 
                 currency = "", 
                 digits = 0, 
                 mark = ",")

```

```{r}

# Get top 10 Latin American countries (excluding Brazil)
combined_data <- COUNTRY_TOP_10 %>%
  filter(
    country_name %in% c("Colombia", "El Salvador", "Honduras", "Nicaragua", 
                       "Venezuela", "Guatemala", "Ecuador", "Paraguay", 
                       "Panama", "Costa Rica", "Peru", "Bolivia", 
                       "Argentina", "Chile") |
    grepl("Dominican Rep", country_name, ignore.case = TRUE)
  ) %>%
  group_by(country_name) %>%
  slice_max(cumulative_weeks_in_top_10, n = 1) %>%
  head(10) %>%
  mutate(season_title = ifelse(is.na(season_title), "N/A", season_title)) %>%
  left_join(
    GLOBAL_TOP_10 %>%
      mutate(season_title = ifelse(is.na(season_title), "N/A", season_title)) %>%
      group_by(show_title, season_title) %>%
      slice_max(week, n = 1),
    by = c("show_title", "season_title"),
    suffix = c("_country", "_global")
  ) %>%
  select(
    Country = country_name,
    `Show Title` = show_title,
    `Season Title` = season_title,
    `Cumulative Weeks` = cumulative_weeks_in_top_10_country,
    `Global Hours Viewed` = weekly_hours_viewed,
    `Global Rank` = weekly_rank_global
  )

# Display first table
datatable(combined_data,
          caption = "Top Latin American Shows: Regional vs Global Performance",
          options = list(pageLength = 10, searching = FALSE, dom = 't', ordering = FALSE),
          rownames = FALSE) %>%
  formatCurrency('Global Hours Viewed', currency = "", digits = 0, mark = ",")

# Get unique shows
show_summary <- combined_data %>%
  filter(!is.na(`Global Hours Viewed`)) %>%
  group_by(`Show Title`, `Season Title`, `Global Hours Viewed`) %>%
  summarize(Countries = n(), .groups = 'drop') %>%
  select(
    show_title = `Show Title`,
    season_title = `Season Title`,
    total_hours = `Global Hours Viewed`,
    countries = Countries
  )

# Add total row
total_row <- tibble(
  show_title = "TOTAL",
  season_title = "",
  total_hours = sum(show_summary$total_hours),
  countries = sum(show_summary$countries)
)

show_summary_final <- bind_rows(show_summary, total_row) %>%
  select(
    `Show Title` = show_title,
    `Season Title` = season_title,
    `Total Hours Viewed` = total_hours,
    `Total Episode Views` = total_hours,
    Countries = countries
  )

# Display summary table
datatable(show_summary_final,
          options = list(pageLength = 10, searching = FALSE, dom = 't', ordering = FALSE),
          rownames = FALSE) %>%
  formatCurrency(c('Total Hours Viewed', 'Total Episode Views'), 
                 currency = "", digits = 0, mark = ",")

```